\chapter{About This Guide}

This Lookup Guide aims to provide a comprehensive summary of all the capabilities of Autodesk Inventor. It has been developed by UQ MARS with the intention of broadening the awareness around CAD software and serving as a manual for students while they they are developing their skills in 3D modelling. It is not intended as a resource to teach students how to use CAD, and it is recommended that students attend the offered CAD Workshop if they are completely new to the softwares. Instead, this is aimed towards students with an intermediate background in CAD and are looken to hone their skillset. This guide will aim to give a broad understanding of Autodek Inventor's various tools and in what applications they can be benefited from, and give greater awareness to powerful tools which are often underutilised. Note that, although this guide is specific to Inventor, most of the tools will also be present in other CAD software.

\section{Using This Guide}
Chapters are organised corresponding to the tabs of the workspace ribbon, and each one will cover a different core capability of Inventor. Over time, content will be added chapter by chapter.
Within each chapter, sections are marked as either Green, Orange or Red. Green sections cover the more fundamental skills which are very important to master, while Orange sections have more niche use cases and Red sections cover advanced topics.

Because the length of this guide will only keep growing, we don't recommend that you try to remember everything that has been written. All key takeaways will be listed at the start of each chapter, and should provide a brief overview of the most important parts of the section. You can then refer to the rest of the chapter if you are interested in a particular feature or if you want to use it for a project and need more information.

We always welcome feedback and will keep updating our chapters for better coverage and readability. If you want a section to be clarified, expanded on or you find an error, you are welcome to provide feedback through Github Issues or via our discord at \url{https://discord.gg/JWzFrKeN}.
